---
title: "Inverse Distributions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Inverse Distributions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(distplyr)
```

An _inverse distribution_ is defined as follows: if $X$ is a random variable with cumulative distribution function (CDF) $F_X$ and $P(X = 0) = 0$, then $Y = 1 / X$ follows an inverse distribution with CDF
$$F_Y(y) = F_X(0) - F_X\left(\frac{1}{y}\right) + P\left(X = \frac{1}{y}\right) + I(y \geq 0).$$

Special cases:

- If $X$ is discrete, then this reduces to $$F_Y(y) = F_X(0) - F_X\left(\frac{1}{y}\right) + f_X\left(\frac{1}{y}\right) + I(y \geq 0),$$ where $f_y$ is the probability mass function. 

- If $X$ is discrete and positive, then this reduces to $$F_Y(y) = 1 - F_X\left(\frac{1}{y}\right) + f_X\left(\frac{1}{y}\right).$$

- If $X$ is continuous, then $$F_Y(y) = F_X(0) - F_X\left(\frac{1}{y}\right) + I(y \geq 0).$$

- If $X$ is continuous and positive, then $$F_Y(y) = 1 - F_X\left(\frac{1}{y}\right).$$

The quantile function is the left inverse of the CDF. Since the $P\left(X = \frac{1}{y}\right)$ portion only applies to a countable set, it does not modify the left inverse. 
